---
title: "Air Pollution Project"
author: "nthehai01"
date: "2/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

For this first programming assignment you will write three functions that are meant to interact with dataset that accompanies this assignment. The dataset is contained in a zip file specdata.zip that you can download from the Coursera web site. 


## Data

#### Downloading data

The zip file containing the data can be downloaded here: [specdata.zip](https://d396qusza40orc.cloudfront.net/rprog%2Fdata%2Fspecdata.zip)

Set up directory for downloading data
```{r}
downloadURL <- "https://d396qusza40orc.cloudfront.net/rprog%2Fdata%2Fspecdata.zip"
downloadedFile <- "./specdata.zip"
```

Download and unzip data
```{r}
if(!file.exists(downloadedFile)) {
    download.file(downloadURL, downloadedFile, method = "curl")
    unzip(downloadedFile)
}
```

#### Data Summary

The zip file contains 332 comma-separated-value (CSV) files containing pollution monitoring data for fine particulate matter (PM) air pollution at 332 locations in the United States. Each file contains data from a single monitor and the ID number for each monitor is contained in the file name. For example, data for monitor 200 is contained in the file "200.csv". Each file contains three variables:

* Date: the date of the observation in YYYY-MM-DD format (year-month-day)
* sulfate: the level of sulfate PM in the air on that date (measured in micrograms per cubic meter)
* nitrate: the level of nitrate PM in the air on that date (measured in micrograms per cubic meter)


## Part 1

Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA.

#### Function summary

*Input:*

* ```directory``` is a character vector of length 1 indicating the location of the CSV files
* ```pollutant``` is a character vector of length 1 indicating the name of the pollutant for which we will calculate the mean; either "sulfate" or "nitrate"
* ```id``` is an integer vector indicating the monitor ID numbers to be used

*Output:*

Return the mean of the pullutant across all monitors list in the ```id``` vector (ignoring NA values)

#### Function definitions

- *Util functions*

```construct.path(id, directory)```: construct the full path from the main directory to a specific given csv file based on the id of the file. Example: if the ```directory``` and ```id``` arguments are ```data``` and ```10```, respectively, the result is ```data/010.csv```.
```{r}
construct.path <- function(id, directory) {
    return(paste(directory, 
                 "/", 
                 formatC(id, width = 3, flag = "0"), 
                 ".csv", 
                 sep = ""))     
}
```
  
```read.monitor(path)```: read the data from a particular monitor stored in a file.
```{r}
read.monitor <- function(id, directory) {
    path <- construct.path(id, directory)
    return(read.csv(path))
}
```

```read.monitors(id, directory)```: read the data from all the monitors and concatenate them
```{r}
library(plyr)
read.monitors <- function(id, directory) {
    ## read all the data frames
    dfs <- lapply(id, read.monitor, directory)  
    
    ## merge data frames
    return(ldply(dfs, rbind))
}
```

- *Main function*
```{r}
pollutantmean <- function(directory, pollutant, id = 1:332) {
    df <- read.monitors(id, directory)  ## load data frames
    return(mean(df[, pollutant], na.rm = TRUE))
}
```

- *Some illustrations*
```{r}
pollutantmean("specdata", "sulfate", 1:10)
```

```{r}
pollutantmean("specdata", "nitrate", 70:72)
```

```{r}
pollutantmean("specdata", "nitrate", 23)
```