---
title: "Air Pollution Project"
author: "nthehai01"
date: "2/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

For this first programming assignment you will write three functions that are meant to interact with dataset that accompanies this assignment. The dataset is contained in a zip file specdata.zip that you can download from the Coursera web site. 


## Data

#### Downloading data

The zip file containing the data can be downloaded here: [specdata.zip](https://d396qusza40orc.cloudfront.net/rprog%2Fdata%2Fspecdata.zip)

Set up directory for downloading data
```{r}
downloadURL <- "https://d396qusza40orc.cloudfront.net/rprog%2Fdata%2Fspecdata.zip"
downloadedFile <- "./specdata.zip"
```

Download and unzip data
```{r}
if(!file.exists(downloadedFile)) {
    download.file(downloadURL, downloadedFile, method = "curl")
    unzip(downloadedFile)
}
```

#### Data Summary

The zip file contains 332 comma-separated-value (CSV) files containing pollution monitoring data for fine particulate matter (PM) air pollution at 332 locations in the United States. Each file contains data from a single monitor and the ID number for each monitor is contained in the file name. For example, data for monitor 200 is contained in the file "200.csv". Each file contains three variables:

* Date: the date of the observation in YYYY-MM-DD format (year-month-day)
* sulfate: the level of sulfate PM in the air on that date (measured in micrograms per cubic meter)
* nitrate: the level of nitrate PM in the air on that date (measured in micrograms per cubic meter)


## Part 1

Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA.

#### Function summary

```pollutantmean(directory, pollutant, id = 1:332)```

*Input:*

* ```directory``` is a character vector of length 1 indicating the location of the CSV files
* ```pollutant``` is a character vector of length 1 indicating the name of the pollutant for which we will calculate the mean; either "sulfate" or "nitrate"
* ```id``` is an integer vector indicating the monitor ID numbers to be used

*Output:*

Return the mean of the pullutant across all monitors list in the ```id``` vector (ignoring NA values)

#### Function definitions

- *Util functions*

```construct.path(id, directory)```: construct the full path from the main directory to a specific given csv file based on the id of the file. Example: if the ```directory``` and ```id``` arguments are ```data``` and ```10```, respectively, the result is ```data/010.csv```.
```{r}
construct.path <- function(id, directory) {
    return(paste(directory, 
                 "/", 
                 formatC(id, width = 3, flag = "0"), 
                 ".csv", 
                 sep = ""))     
}
```
  
```read.monitor(path)```: read the data from a particular monitor stored in a file.
```{r}
read.monitor <- function(id, directory) {
    path <- construct.path(id, directory)
    return(read.csv(path))
}
```

- *Main function*
```{r}
library(plyr)
pollutantmean <- function(directory, pollutant, id = 1:332) {
    ## read all the data frames
    dfs <- lapply(id, read.monitor, directory)  
    
    ## concatenate data frames in dfs
    df <- ldply(dfs, rbind)
    
    return(mean(df[, pollutant], na.rm = TRUE))
}
```

- *Some illustrations*
```{r}
pollutantmean("specdata", "sulfate", 1:10)
```

```{r}
pollutantmean("specdata", "nitrate", 70:72)
```

```{r}
pollutantmean("specdata", "nitrate", 23)
```


## Part 2

Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases.

#### Function summary

```complete(directory, pollutant, id = 1:332)```

*Input*

* ```directory``` is a character vector of length 1 indicating the location of the CSV files
* ```id``` is an integer vector indicating the monitor ID numbers to be used

*Output*

Return a data frame of the form:
```{r}
##    id      nobs
##    1       117
##    2       1041
##    ...
```

where ```id``` if the monitor ID number and ```nobs``` is the number of complete cases.

#### Function definition

```{r}
complete <- function(directory, id = 1:332) {
    ## read all the data frames
    dfs <- lapply(id, read.monitor, directory)  
    
    ## check whether an observation in dfs is complete
    cpte.cases.rough <- lapply(dfs, complete.cases)
    ## sum of completely observed cases
    cpte.cases.rough <- sapply(cpte.cases.rough, sum)
    
    ## return the result as a proper data frame
    return(data.frame(
        id = id,
        nobs = cpte.cases.rough
    ))
}
```

- *Some illutrations*

```{r}
complete("specdata", 1)
```

```{r}
complete("specdata", c(2, 4, 8, 10, 12))
```

```{r}
complete("specdata", 30:25)
```

```{r}
complete("specdata", 3)
```


## Part 3

Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0.

#### Function summary

```corr(directory, threshold = 0)```

*Input*

* ```directory``` is a character vector of length 1 indicating the location of the CSV files
* ```threshold``` is a numeric vector of length 1 indicating the number of completely observed observation (on all variables) required to compute the correlation between nitrate and sulfate; the default is 0

*Output*

Return a numeric vector of correlations

#### Function definitions

- *Util function*

```get.cor(df)```: get the correlation between ```sulfate``` and ```nitrate``` in ```df```.
```{r}
get.cor <- function(df) cor(df$sulfate, df$nitrate, use = "complete.obs")
```

- *Main function*

```{r}
corr <- function(directory, threshold = 0) {
    ## find the number of completely observed cases on each monitor
    cpte.cases <- complete(directory)
    
    ## filter the monitors that meet the threshold
    proper.cases <- cpte.cases[cpte.cases$nobs > threshold, ]
    
    ## read data of the monitors that meets the threshold
    dfs <- lapply(proper.cases$id, read.monitor, directory)  
    
    return(sapply(dfs, get.cor))
}
```

- *Some illustrations*

```{r}
cr <- corr("specdata", 150)
head(cr)
```

```{r}
summary(cr)
```

```{r}
cr <- corr("specdata", 400)
head(cr)
```

```{r}
summary(cr)
```

```{r}
cr <- corr("specdata", 5000)
summary(cr)
```

```{r}
length(cr)
```

```{r}
cr <- corr("specdata")
summary(cr)
```

```{r}
length(cr)
```

